---
title: "Midterm 2018"
author: "C. Durso"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(stringr)
library(MASS)
```
Name_____________________________________

## Midterm Exam, COMP 4441, 3441: Probability and Statistics for Data Science

###Due Friday, February 16, 11:59pm

This is an open book, open note exam. You may use any online or printed reference materials, but you are not expected to go beyond what can be deduced in a basic manner from the text and course notes. You may not use responses to questions you pose to other people, online or directly. 
You may ask the instructor for clarification of questions. Do not ask anyone other than the instructor anything relating to the exam. Using anyone else's help on the exam is a violation of the conditions of the exam, and is cause for action including, but not limited to, a 0 on the exam. It is also a breach of academic trust. (Sorry to sound so nasty, but having this clear is important.)

Please complete the following collection of descriptions of procedures. The first entry has been completed as an example. Each question is worth 20 points.


#### 1. Procedure name: Sign test

Data requirements [one, two, or more samples, data type (numeric, ordinal, or categorical), data distribution (Normal, symmetric, non-parametric, etc.), sample size?]: 

As discussed in class, the sign test applies to a single sample of size $n$ of numeric data with positive and negative values.

Formulas for the statistic(s), and p-value, if relevant:

The statistic is the number $k$ of positive values in the sample, considered as a draw from a binomial distribution with size equal to $n$ and probability of success (positive value) equal to $0.5$.

The p-value for the two-sided test is $\sum_{m:f(m,n,0.5)\leq f(k,m,0.5)}f(m,n,0.5)$, the sum of all probabilities of all outcomes with probability less than or equal to that of the observed number of positive values under the binomial distribution with size equal to $n$ and probability of success equal to $0.5$.

Interpretation (Is there a null hypothesis? Associated confidence interval?) :

The null hypothesis is that each observation in the sample is equally likely to be positive or non-positve. If there is neglible probability that an observation equals 0, this is equivalent to the hypothesis that the median of the distribution is equal to 0. No confidence interval was mentioned. 


#### 2. Procedure name: One sample z-test

Data requirements [one, two, or more samples, data type (numeric, ordinal, or categorical), data distribution (Normal, symmetric, non-parametric, etc.), sample size?]: 

Formulas for the statistic(s), and p-value, if relevant: 

Interpretation (Is there a null hypothesis? Associated confidence interval?) :

#### 3. Procedure name: One sample Wilcoxon signed rank test

Data requirements [one, two, or more samples, data type (numeric, ordinal, or categorical), data distribution (Normal, symmetric, non-parametric, etc.), sample size?]:

Formulas for the statistic(s), and p-value, if relevant:

Interpretation (Is there a null hypothesis? Associated confidence interval?) : 

#### 4. Procedure name: One sample Student's t-test

Data requirements [one, two, or more samples, data type (numeric, ordinal, or categorical), data distribution (Normal, symmetric, non-parametric, etc.), sample size?]:

Formulas for the statistic(s), and p-value, if relevant:

Interpretation (Is there a null hypothesis? Associated confidence interval?) : 

#### 5. Procedure name: bootstrap test for the value of the mean

Data requirements [one, two, or more samples, data type (numeric, ordinal, or categorical), data distribution (Normal, symmetric, non-parametric, etc.), sample size?]:

Formulas for the statistic(s), and p-value, if relevant:

Interpretation (Is there a null hypothesis? Associated confidence interval?) :



#### 6. Applications
In R, apply each of the methods from questions 2-5 to each of the data vectors v1, v2, and v3 loaded below, using the null hypothesis $\mu_0=0$, and $\sigma=1$ if necessary. Report the p-value, and a confidence interval for the mean, if readily available. State whether or not the method is appropriate for the data.


```{r}
load("midterm_data.RData")
#Use QQ-Plots to check for normality:

#Not Normal
qqnorm(v1)
qqline(v1)

#Normal
qqnorm(v2)
qqline(v2)

#Eh, Vaguely normal
qqnorm(v3)
qqline(v3)


framev1 <- data.frame(v1)
framev2 <- data.frame(v2)
framev3 <- data.frame(v3)

gV1 <- ggplot(framev1)+geom_histogram(aes(x = v1, y = ..density..))+
  stat_function(fun = dnorm, n = length(v1), args = list(mean = mean(v1), sd = sd(v1)), color = "red")
gV1

gV2 <- ggplot(framev2)+geom_histogram(aes(x = v2, y = ..density..))+
  stat_function(fun = dnorm, n = length(v2), args = list(mean = mean(v2), sd = sd(v2)), color = "red")
gV2

gV3 <- ggplot(framev3)+geom_histogram(aes(x = v3, y = ..density..))+
  stat_function(fun = dnorm, n = length(v3), args = list(mean = mean(v3), sd = sd(v3)), color = "red")
gV3


```


  + 6.a. One Sample z-test:
```{r} 
#Data Vector V1:
nVec1 = nrow(v1)
sampleSigma = 1
SEMeanV1 = sampleSigma/sqrt(nVec1)
EV1 = qnorm(.975)*SEMeanV1; #Margin of Error
xbarV1 = mean(v1)
cat("Mean V1: ",xbarV1, "\n")
cat("Confidence Interval V1: ", xbarV1 + c(-EV1, EV1), "\n")

#Is m = 0?
zV1 <- (xbarV1-0)/SEMeanV1
pV1 <- 2*pnorm(-abs(zV1))
cat("P-Value for V1: ", pV1, "\n\n")

library(TeachingDemos)
z.test(v1, sd = sampleSigma)

#Data Vector V2:
nVec2 = length(v2)
SEMeanV2 = sampleSigma/sqrt(nVec2)
EV2 = qnorm(.975)*SEMeanV2 #Margin of Error
xbarV2 = mean(v2)
cat("\nMean V2: ",xbarV2, "\n")
cat("Confidence Interval V2: ", xbarV2 + c(-EV2, EV2), "\n")
zV2 <- (xbarV2-0)/SEMeanV2
pV2 <- 2*pnorm(-abs(zV2))
cat("P-Value for V2: ", pV2, "\n\n")

z.test(v2, sd = sampleSigma)


#Data Vector V3:
nVec3 = nrow(v3)
SEMeanV3 = sampleSigma/sqrt(nVec3)
EV3 = qnorm(.975)*SEMeanV3 #Margin of Error
xbarV3 = mean(v3)
cat("\nMean V3: ",xbarV3, "\n")
cat("Confidence Interval V3: ", xbarV3 + c(-EV3, EV3), "\n")
zV3 <- (xbarV3-0)/SEMeanV3
pV3 <- 2*pnorm(-abs(zV3))
cat("P-Value for V3: ", pV3, "\n\n")
z.test(v3, sd = sampleSigma)

``` 
#Response to 6.A:
Based on the p-values returned for all three data vectors, we can reject the null hypothesis that the mean of the vectors is equal to zero for V1 and V3, as the p-values for this test were less than or equal to 0.05 for the V1 and V3 datasets.For V2, the p-value for this test was greater than .05, as such we cannot reject the null hypothesis based on this particular test. 

Given that the Z-Test expects the data to follow a normal distribution, I plotted each vector using qqnorm/qqline to check for normality of the data sets.Based on their qqplots, V2 appears to follows a normal distibution, V3 vaguely follows a normal distribution, and V1 does not follow a normal distribution. Given this, the one-sample Z-Test is a good statistical test for data vector V2, and has potential to be a good statistical test for data vector V3 (although there may be an alternative statistical test that would better fit V3). However, the one-sample Z-Test is not a good statistical test for data vector V1 based on the distribution of the data.


  + 6.b. Wilcoxon signed rank test
```{r}

#Data Vector V1:
#summary(v1)

yV1 <- v1-median(v1)
#cat("True Median of V1: ", median(v1), "\n")
yrankV1 <- rank(abs(yV1))
cat("Value of V for V1: ",sum(yrankV1[yV1>0]),"\n")
wilcox.test(v1, mu = median(v1))

#Data Vector V2 (does look normal):
#summary(v2)

yV2 <- v2-median(v2)
yrankV2 <- rank(abs(yV2))
#cat("True Median of V2: ", median(v2), "\n")
cat("\n\nValue of V for V2: ",sum(yrankV2[yV2>0]),"\n")
wilcox.test(v2, mu = median(v2))

#Data Vector V3:
#summary(v3)
yV3 <- v3-median(v3)
yrankV3 <- rank(abs(yV3))
#cat("True Median of V3: ", median(v3), "\n")
cat("\n\nValue of V for V3: ",sum(yrankV3[yV3>0]),"\n")
wilcox.test(v3, mu = median(v3))


```
#Response to 6.B:
Based on the p-values returned for all three data vectors, we cannot reject the null hypothesis for the there datasets as all three p-values were greater than or equal to 0.05. Given that the Wilcoxon Signed Mean Test expects that the given data will follow a non-normal distribution, whether or not this statistical test is a good fit for the given data will depend on the normality of the datasets. 

Based on the qqplots discussed in 6.A, the Wilcoxon Signed Rank Test is a good statistical test for the data vector V1, and has potential to be a good statistical test for data vector V3 (although there may be an alternative statistical test that would better fit V3). However, the Wilcoxon Signed Rank Test is not a good statistical test for data vector V2 based on the normal distribution of the data.

  + 6.c. Student's t-test TODO: Is this a good test?
```{r}
#Data Vector V1:

#Confidence Interval:
nV1 <- nrow(v1)
nullHypMean <- 0 #Using Null Hypothesis Mean
sampleSDV1 <- sd(v1)
sampleMeanV1 <- mean(v1)
SEV1 = sampleSDV1/sqrt(nV1)
EV1 = qt(.975, df = nV1-1)*SEV1
cat("Confidence Interval for V1: ", sampleMeanV1+c(-EV1, EV1), "\n")

#P-Value:
tTestV1 <- (nullHypMean-sampleMeanV1)/SEV1
cat("P-Value for V1: ", 2*pt(tTestV1,nV1-1), "\n\n")
t.test(v1, mu = nullHypMean, conf.level = 0.95)


#Data Vector V2:

#Confidence Interval:
nV2 <- length(v2)
sampleSDV2 <- sd(v2)
sampleMeanV2 <- mean(v2)
SEV2 = sampleSDV2/sqrt(nV2)
EV2 = qt(.975, df = nV2-1)*SEV2
cat("\nConfidence Interval for V2: ", sampleMeanV2+c(-EV2, EV2), "\n")

#P-Value:
tTestV2 <- (nullHypMean-sampleMeanV2)/SEV2
cat("P-Value for V2: ", 2*pt(tTestV2,nV2-1), "\n\n")
t.test(v2, mu = nullHypMean, conf.level = 0.95)

#Data Vector v3:

#Confidence Interval:
nV3 <- nrow(v3)
sampleSDV3 <- sd(v3)
sampleMeanV3 <- mean(v3)
SEV3 = sampleSDV3/sqrt(nV3)
EV3 = qt(.975, df = nV2-1)*SEV3
cat("\nConfidence Interval for V3: ", sampleMeanV3+c(-EV3, EV3), "\n")

#P-Value:
tTestV3 <- (nullHypMean-sampleMeanV3)/SEV3
cat("P-Value for V3: ", 2*pt(tTestV3,nV3-1), "\n\n")
t.test(v3, mu = nullHypMean, conf.level = 0.95)
```
#Response to 6.C:
Based on the p-values returned for all three data vectors, we can reject the null hypothesis that the mean of the vectors is equal to zero for all datasets, as the p-values returned from this test were less than or equal to 0.05.

Based on the qqplots discussed in 6.A, and given that the T-Test expects the data to follow a normal distribution, this statistical test is a good fit for data vector V2 and has potential to be a good statistical test for data vector V3 (although there may be an alternative statistical test that would better fit V3). However, the one-sample T-Test is not a good statistical test for data vector V1 based on the non-parametric distribution of the data.

  + 6.d. bootstrap test for the value of the mean #To Do: Is this a good test?
```{r}
nSamples <- 10000
grabMeans <- function(x){
  sampleMean <- mean(x)
  return (sampleMean)
  
}

#Data Vector V1:
listSamplesV1 <- lapply(1:nSamples, function(x) sample(v1, size = nrow(v1), replace = TRUE))
meanListV1 <- sapply(listSamplesV1, grabMeans)

#Used to ensure normality:
#qqnorm(meanListV1)
#qqline(meanListV1)

cat("Bootstrapped interval V1: ", quantile(meanListV1, c(0.025, 0.975)), "\n")
cat("Sample mean of the list of V1 means: ", mean(meanListV1), "\n")
cat("Sample mean of V1: ", mean(v1), "\n\n")

#Data Vector V2:
listSamplesV2 <- lapply(1:nSamples, function(x) sample(v2, size = length(v2), replace = TRUE))
meanListV2 <- sapply(listSamplesV2, grabMeans)

#qqnorm(meanListV2)
#qqline(meanListV2)

cat("Bootstrapped interval V2: ", quantile(meanListV2, c(0.025, 0.975)), "\n")
cat("Sample mean of the list of V2 means: ", mean(meanListV2), "\n")
cat("Sample mean of V2: ", mean(v2), "\n\n")


#Data Vector V3:
listSamplesV3 <- lapply(1:nSamples, function(x) sample(v3, size = nrow(v3), replace = TRUE))
meanListV3 <- sapply(listSamplesV3, grabMeans)

#qqnorm(meanListV3)
#qqline(meanListV3)

cat("Bootstrapped interval V3: ", quantile(meanListV3, c(0.025, 0.975)), "\n")
cat("Sample mean of the list of V3 means: ", mean(meanListV3), "\n")
cat("Sample mean of V3: ", mean(v3), "\n\n")


```
#Response to 6.D:
Based on the 95% bootstrapped mean intervals returned for all three data vectors, the Bootstrapped Interval for the Mean test appears to be a good test for the mean of all three datasets.



