---
title: "Problem Set 2"
author: "Ariel Huckabay"
date: "`r format(Sys.time(),'%B %d, %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
These questions were rendered in R markdown through RStudio (<https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>, <http://rmarkdown.rstudio.com> ).

Please complete the following tasks regarding the data in R. Please generate a solution document in R markdown and upload the .Rmd document and a rendered  .doc, .docx, or .pdf document.  Your work should be based  on the data's being in the same folder as the .Rmd file. Please turn in your work on Canvas. Your solution document should have your answers to the questions and should display the requested plots.


For use in the problems 1, 2, 3, and 4 below, generate a vector of 100,000 values from the binomial distribution with n=20 and p=0.5 using the seed 7654321. 

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(knitr)

set.seed(7654321)
randBinom <- rbinom(100000, 20, .5)
randVector<-vector("numeric", 100000)
randVector = randBinom

```

1. Present a histogram of the results using the techniques in distributions.Rmd.(10 points) 

```{r}
hist(randVector)

```

2. Discuss the appearance of the histogram in relation to the probability density function of the binomial distribution with n=20 and p=.5. (10 points)



```{r}
rMean <- mean(randVector)
stdDev<-sd(randVector)
```

The histogram approximates a normal distribution (bell-shaped curve). Half of the values fall beyond the mean of `r rMean`, and half fall beore it by visual inspection.  

3. Create a plot with the histogram from above with the density of the normal distribution with mean equal to 20 and sigma equal sqrt(50\*.4\*(1-.4)). Do this binomial distribution and this normal distribution appear related?  (10 points)

```{r}
NormPlot= rnorm(100000, 20, sqrt(50 *0.4*(1-0.4)))
hist(NormPlot)
```

Both of these historgrams have a bell-curve shape, but this normal distribution is steeper due to a standard deviation that is lower relative to the mean.  


4. Now consider the probability space generated by applying the random variable X with X(y)=(y-50\*.4)/sqrt(50\*.4\*(1-.4)) to the Binomial distribution with size=50 and probability=.4 Please plot the density of this distribution. Also, setting d=1/sqrt(50\*.4\*(1-.4)), plot the points with x-values equal to the outcomes of the new distribution and y-values equal to the probability that a sample from the standard normal distribution lies between x-d/2 and x+d/2 for each of these x's. Do the two distributions seem to be closely related? (10 points) 

```{r}
#set d
d=1/sqrt(50*0.4*(1-0.4))

#random variable 
X<-function(x){(x-50*0.4)/(sqrt(50*0.4*(1-0.4)))}

#filtering function to get what's equal to d
findD<-function(x){if (x == d) return(TRUE) else return(FALSE)}

#initialize the y's and the vector of x's applying the specified random variable function
vecY<-rbinom(100000, 50, 0.4)
vecX<-sapply(vecY, X)
frame<-data.frame(vecX)

dStruct<-Filter(findD, vecX)


hist(vecX)


```


5. For some values of size s and probability p, the standard normal distribution is a good approximation to the distribution that arises from applying the random variable X defined by X(y)=(y-s\*p)/sqrt(s\*p\*(1-p)) to the binomial distribution with size s and probability p. For which of the following pairs (s,p): (50,.5), (50,.1), (10,.1), (50,.01),(5000,.01) does that seem to be the case? Please support your conclusion with visualizations such as the one in question 4. You may use density functions without any sampling. You may find it helpful to extend the x-values for the normal distribution to be symmetric around zero using z<-unique(c(x,-x)), for example. Also, you can control the limits of the x-axis with g<-g+xlim(-3,3), say (10 points)

```{r}

```

6. (4441 only) Recall the normal probabilities from question 4: the x values are (z-50\*.4)/sqrt(50\*.4\*(1-.4)) for the integers z in {0,1,2,3,...50}. The value d is defined by d=1/sqrt(50\*.4\*(1-.4)), For each x value, the point (x,y) where y is probability that a draw from the standard normal distribution lies between x-d/2 and x+d/2 was plotted. If you plot the density of the standard normal distribution on the same plot, it appears to be of a similar shape but different vertical scale. What can you multiply the standard normal density function by to make the scales comparable? Please give a formula in terms of s and p and explain your reasoning.(10 points)

```{r}

```


