---
title: "Problem Set 5"
author: "cdurso"
date: "Thursday, January 21, 2016"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

These questions were rendered in R markdown through RStudio (<https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>, <http://rmarkdown.rstudio.com> ).

Please complete the following tasks, in R where applicable. Please generate a solution document in R markdown and upload the rendered .doc, .docx, or .pdf document.  You may add hand computations to a .doc or .docx if you prefer. In the rendered document, please show your code. That is, don't use "echo=FALSE".  

In either case, your work should be based  on the data's being in the same folder as the R files. Please turn in your work on Canvas. Your solution document should have your answers to the questions and should display the requested plots.

 1. Let $Y$ be the random variable $X^{2}$ where $X$ has a standard Normal distribution $N(0,1)$. Calculate the mean of $Y$. You may use what you know about the mean and variance of $X$. (10 points)

The variance of $X$ equals $E[X^2]-E[X]^2$. Given that $Var[X]=1$ and $E[X]=0$, conclude $E[Y]=E[X^2]=1$.


 2. Returning to the data from "precipitation_boulder.csv", the file "precipitation.RData" contains the data.frame dat.trim created from the code in ps4, followed by the command 

save(dat.trim,file="precipitation.RData")

The command

load("precipitation.RData")

imports dat.trim directly into the current environment.

  * 2.a. Are the annual precipitation totals in 1900-1915 approximately Normally distributed? Are the annual precipitation totals in 2000-2010 approximately Normally distributed? Please provide visual support for your conclusions. (10 points)
  
```{r}
library(tidyverse)
load("precipitation.RData")

#retrieve the 1900-1915 years
dat.1915<-filter(dat.trim, dat.trim$year <= 1915)
dat.1915<-filter(dat.1915, dat.1915$year >= 1900)

#retrieve the 2000-2010
dat.2010<-filter(dat.trim, dat.trim$year <= 2010)
dat.2010<-filter(dat.2010, dat.2010$year >= 2000)

qqnorm(dat.1915$year.total)
qqline(dat.1915$year.total)

qqnorm(dat.2010$year.total)
qqline(dat.2010$year.total)



```

  * 2.b. Perform Welch's t test to test the hypothesis that the annual precipitation totals in 1900-1915 and the annual precipitation totals in 2000-2010 are drawn from populations with equal means. Please comment on the applicability of the test, based on your observations in 2.a. (5 points)
  
```{r}

t.test(dat.1915$year.total,dat.2010$year.total)

```

  * 2.c. Are the variances of the annual precipitation totals in 1900-1915 and the annual precipitation totals in 2000-2015 approximately equal? (5 points)
  
```{r}

#The question asks for 2000-2015 rather than 2000-2010, something to clarify with Cathy
#Is it really this easy?

(var(dat.1915$year.total))
(mean(dat.1915$year.total))

(var(dat.2010$year.total))
(mean(dat.2010$year.total))
```



  * 2.d. Test whether the April precipitation in the years 2000-2010 may reasonably be viewed as coming from a population with same mean as the corresponding June precipitation. (10 points)

```{r}

aprilPrecip <- dat.2010$apr
junePrecip <- dat.2010$jun

aFrame <- data.frame(aprilPrecip)
jFrame <- data.frame(junePrecip)

g1 <- ggplot(aFrame, aes(x=dat.2010$year, y = aprilPrecip)) +geom_point(color="red")
g2 <- g1+geom_point(data = jFrame, aes(x=dat.2010$year, y = junePrecip))
g2 



```

 3. Please describe your project. If you are explaining a method not covered in class, what is the method? If you are addressing a question on a new data set, what is the data set? If you are working in a team of two, who is your teammate? (10 points)





