---
title: "Problem Set 7"
author: "cdurso"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

These questions were rendered in R markdown through RStudio (<https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>, <http://rmarkdown.rstudio.com> ).

Please complete the following tasks, in R where applicable. Please generate a solution document in R markdown and upload the rendered .doc, .docx, or .pdf document.  You may add hand computations to a .doc or .docx if you prefer. In the rendered document, please show your code. That is, don't use "echo=FALSE". 

In either case, your work should be based  on the data's being in the same folder as the R files. Please turn in your work on Canvas. Your solution document should have your answers to the questions and should display the requested plots.

1. Suppose the $Y_i=mX_i+b+\varepsilon_i$ where the $\varepsilon_i$'s are independent, identically distributed $Normal\left(0,\sigma^2\right)$ random variables. Define $M$ and $B$ as in the notes. What is the expected value of $Y_i-\left(X_iM+B\right)$ ?(10 points)


2.	(1 point each) In the following, please use the"ads.txt"data, obtained from http://lib.stat.cmu.edu/DASL/Datafiles/tvadsdat.html then reformatted to load correctly using "read.table" with the option "header=TRUE". The data description follows: "This data appeared in the Wall Street Journal. The advertisement were selected by an annual survey conducted by Video Board Tests, Inc., a New York ad-testing company, based on interviews with 20,000 adults who were asked to name the most outstanding TV commercial they had seen, noticed, and liked. The retained impressions were based on a survey of 4,000 adults, in which regular product users were asked to cite a commercial they had seen for that product category in the past week.

FIRM: Firm name 

SPEND: TV advertising budget, 1983 ($ millions)

MILIMP: Millions of retained impressions per week"

```{r}
library(tidyverse)
dataAds <- read.table("ads.txt", header = TRUE)
dataAds

```
  - 2.a.	Use maximum likelihood regression to model MILIMP as a linear function of SPEND. Provide any diagnostic plots you feel are relevant. Is the relationship linear? Is the assumption of independent, identically distributed Normal errors justifiable? (15 points)
  
```{r}
g <- ggplot(data=dataAds, aes(x=dataAds$MILIMP, y = dataAds$SPEND))+geom_point()
g

milimp <- dataAds$MILIMP
spend <- dataAds$SPEND

n <- nrow(dataAds)
n
m <- (sum(milimp*spend)/n-mean(milimp)*mean(spend))/(sum((milimp)^2)/n-mean(milimp^2))
b <- mean(spend)-m*mean(milimp)
g1 <- g+geom_abline(slope = m, intercept = b)
g1

lm.milimp<-lm(spend~milimp)

spendPredict <- m*milimp+b
spendPredict
predict(lm.milimp)
lm.milimp$fitted

#milimp "~ = is modeled as" linear function of spend
lmfit <- lm(milimp~spend)
lmfit
summary(lmfit)
plot(lmfit)


#This link is a good: https://datascienceplus.com/correlation-and-linear-regression/

#Is the relationship linear?
  #Initially by looking at graphs 1 and 3 , it looks like this model 
  #doesn't quite fit


#Is the assumption of independent, identically dist. normal errors justifiable?

```

  - 2.b.	Use maximum likelihood regression to model  log(MILIMP) as a linear function of log(SPEND). Provide any diagnostic plots you feel are relevant. Is the relationship linear? Is the assumption of independent, identically distributed Normal errors justifiable? (15 points) 

```{r}
logM <- log(milimp)
logS <- log(spend)

LogLmFit <- lm(logM~logS)
LogLmFit
summary(LogLmFit)
plot(LogLmFit)

#Is the relationship linear?
  #Initially by looking at graphs 1 and 3 , it looks like this model fits


#Is the assumption of independent, identically dist. normal errors justifiable?

```

 

  - 2.c Which of the two regressions above looks like a better fit to the data? (5 points)
  
  
  - 2.d For the model you chose in 2.c, what is the predicted MILIMP for SPEND=75? (5 points)
  
```{r}

```


